---
title: "Assignment 2"
author: "LI Yixuan"
date: "07/10/2017"
output: html_document
---
<br/>
```{r setup, include=FALSE}
source("functions.R")
```
<br/>
<br/>
<br/>

## Exercise 1: Writing R functions
<br/>

### a. Filling in the blanks
<br/>
```{r}
print(sum_column(iris, "Sepal.Length"))
print(sum_column(iris, "Species"))
print(sum_column(warpbreaks, "breaks"))
```
<br/>

### b. A function with one argument and one return value
<br/>
```{r}
print(my_sum(iris$Sepal.Length))
print(my_sum(iris$Species))
print(my_sum(warpbreaks$breaks))
```
<br/>

### c. A function with two arguments and one return value
<br/>
```{r}
print(sum_divided_by(iris$Sepal.Length, 12))
print(sum_divided_by(iris$Species, 22))
print(sum_divided_by(iris$Sepal.Length, "Not numeric"))
print(sum_divided_by(warpbreaks$breaks, -12))
```
<br/>

### d. One more function
<br/>
```{r}
print(my_mean(iris$Sepal.Length))
print(my_mean(iris$Species))
print(my_mean(warpbreaks$breaks))
```
<br/>
<br/>
<br/>
<br/>
<br/>

## Exercise 2: Working with ggplot
<br/>

### a. Creating violin plots
<br/>
```{r}
print(grouped_violin_plot(iris, "Sepal.Length", "Species"))
```
<br/>

### b. Modifying plots
<br/>
```{r}
p <- grouped_violin_plot(iris, "Sepal.Length", "Species")
# YOUR CODE HERE: Change the colour scheme for the interior of the three violin plots
#                 to anything else at all.
p <- p + ggplot2::scale_fill_brewer(palette="Blues")
# YOUR CODE HERE: Add a main title that says "Iris data".
p <- p + ggplot2::labs(title="Iris data")# + ggplot2::theme_classic()
print(p)
```
<br/>
<br/>
<br/>
<br/>
<br/>

## Exercise 3: Permutation tests
<br/>

### a. Writing a more generic function for taking a test statistic
<br/>
```{r}
difference_in_medians(iris, "Sepal.Width", "Species", "versicolor", "virginica")
difference_in_medians(iris, "Sepal.Width", "Species", "virginica", "virginica")
```
<br/>

### b. Writing a more generic function of the randomization function
<br/>
```{r}
iris$Sepal.Width[1:10]
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)

randomize(iris, "Sepal.Width")$Sepal.Width[1:10]
randomize(iris, "Species")$Species[1:10]
randomize(iris, "Species")$Sepal.Width[1:10]

set.seed(previous_seed)
```
<br/>

### c. A function to get a statistic for multiple permutations
<br/>
```{r}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)

ptest_1 <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor","virginica", difference_in_medians, n_samples=10)
ptest_2 <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor","virginica", difference_in_medians, n_samples=10)
ptest_3 <- permutation_twogroups(randomize(iris, "Sepal.Width"), "Sepal.Width","Species", "versicolor", "virginica", difference_in_medians, n_samples=10)
set.seed(previous_seed)

print(ptest_1)
print(ptest_2)
print(ptest_3)
print(ptest_3$observed)
print(ptest_3[["observed"]])
```
<br/>

### d. Does it matter for your permutation test whether you permute var or grouping_var? 
<br/>
It matters, because as we have defined the permutation function like `d[[var]] <- sample(d[[var]], nrow(d))`, it will only change the order of elements in the column indicated by `var`. If we use the `grouping_var` instead, the data not in the column "Sepal.Width", but in the column "Condition" will get permuted.<br/>
Therefore, when we change the argument of *randomize* function from *var* to *grouping_var*, the difference can be seen in the result numbers. And if we do so, maybe the permutation won't be as well as that using "Sepal.Width".<br/>
<br/>
But this kind of difference in result here seems won't obstruct us much from drawing a correct conclusion. Wheather permute var or grouping_var, as long as the sample are large enouugh, the distributions are quite close. <br/>
<br/>

### e. Plotting the sampling distribution
<br/>
```{r cache=TRUE}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor", "virginica", difference_in_medians)
set.seed(previous_seed)
```
<br/>
```{r}
ptest_d <- tibble::as_tibble(ptest["permuted"])
#print(ptest_d)

# YOUR CODE HERE: plot a histogram with a vertical line at the observed value
ggplot2::ggplot(ptest_d, ggplot2::aes(x=`permuted`)) +
ggplot2::geom_histogram(bins=30, fill="grey", colour="black") + ggplot2::geom_vline(ggplot2::aes(xintercept=difference_in_medians(iris, "Sepal.Width", "Species", "versicolor","virginica")), lwd=0.5, color="green")
```
<br/><br/>
In the graph, it's clear eo see the difference between 2 the two datasets : at -0.2 is the difference of mediens of each group in the original iris dataset, which means between the two species of iris, their "Sepal Width" differ generally 0.2cm; while, with distribution presented by histogram, this difference in the permuted data set have both its mode and medien at 0.0, which means the two groups can be considered as the same.
<br/>
Therefore, we can say that in the original data set, the two species are not the same in "Sepal Width".<br/>
<br/>

### f. Test statistics
<br/>
```{r}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest_new <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor", "virginica", new_test_statistic)
set.seed(previous_seed)
```
<br/>
```{r}
ggplot2::ggplot(ptest_d,ggplot2::aes(x=`permuted`)) + 
  ggplot2::geom_histogram(fill="grey",color="black") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=new_test_statistic(iris, "Sepal.Width", "Species", "versicolor","virginica")), lwd=0.5, color="yellow")
```
<br/><br/>
Here we construct a new test statistic function who calculate the difference in means of two groups of flowers. The new line and distribution histogram dose not differ much from those of difference in mediens: only the line move slightly to the left. <br/>
This can be explained by the fact that the distribution of data examples could lead to a difference between the medien and mean.<br/>
In general, the two test statistic functions give us a similar anwser supporting the *Hypothesis B*. <br/>
<br/>

### g. Calculating p-values
<br/>
```{r}
permutation_pvalue_right(ptest)
permutation_pvalue_left(ptest)
```
<br/><br/>
Let's take the example of *the height of a group of high school studients, where there are girls and boys*.<br/>
<br/>
We'd predict that the difference in the medians statistic would be extreme on the right
side if our **Hypothesis B** were that *the height of studients of different gender follows different distribution : boys are usually higher than girls. So for calculating the difference in mediens, we subtract the boys' by the girls'*.<br/>
<br/>
In which case, we would use the function `permutation_pvalue_right`, because when the value is extreme on the right (in another word, in the positive direction), we will observe a extremely small number as result of this function.<br/>
<br/>

### h. Writing a more generic function of the randomization function
<br/>
```{r cache=TRUE}
iris_subset1 <- iris[c(89:94, 108:112),]
iris_subset2 <- iris[88:114,]

if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest1 <- permutation_twogroups(iris_subset1, "Sepal.Width", "Species", "versicolor", "virginica", difference_in_medians)
ptest2 <- permutation_twogroups(iris_subset2, "Sepal.Width", "Species", "versicolor", "virginica", difference_in_medians)
set.seed(previous_seed)
```
<br/>
```{r cache=TRUE}
ptest_d1 <- tibble::as_tibble(ptest1["permuted"])
ptest_d2 <- tibble::as_tibble(ptest2["permuted"])

ggplot2::ggplot(ptest_d1,ggplot2::aes(x=`permuted`)) + 
  ggplot2::geom_histogram(fill="darkblue",color="black") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=new_test_statistic(iris, "Sepal.Width", "Species", "versicolor","virginica")), lwd=0.5, color="green") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=new_test_statistic(iris_subset1, "Sepal.Width", "Species", "versicolor","virginica")), lwd=0.5, color="blue") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=new_test_statistic(iris_subset2, "Sepal.Width", "Species", "versicolor","virginica")), lwd=0.5, color="red") + ggplot2::xlim(c(-0.6,0.6))

ggplot2::ggplot(ptest_d2,ggplot2::aes(x=`permuted`)) + 
  ggplot2::geom_histogram(fill="darkred",color="black") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=new_test_statistic(iris, "Sepal.Width", "Species", "versicolor","virginica")), lwd=0.5, color="green") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=new_test_statistic(iris_subset1, "Sepal.Width", "Species", "versicolor","virginica")), lwd=0.5, color="blue") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=new_test_statistic(iris_subset2, "Sepal.Width", "Species", "versicolor","virginica")), lwd=0.5, color="red") + ggplot2::xlim(c(-0.6,0.6))
```
<br/><br/>
It can be noticed that in both graph, the defference between two species become large (with less data around 0.0) and a lot more concentrate, as we select some data of a certain range of the original dataset instead of it all. With a central part missing, while remained bars of the second subdataset keep a similar shape with the origianl one, those of the first subdataset distribute in a more  dispersed way on the x axis.<br/>
<br/>
A second notation is in regard to the position of vertical lines, which all at the left side of the whole data points tell that the subtraction would more like to get a negtive value before permutation. The full dataset and its subdataset2 have a close value, but become quite extreme with the subdataset 2 which has selected data form rows between from 89th to 94th and from 108th to 112th. As the range is smaller, it concains less data values thus can lead to a deviation.<br/>
<br/>
After all permutation test above we can be pretty sure that these data are belong to two sepecies of iris, rather than a sigular. <br/>
For the level of etremeness, although the “fairly extreme” line with subdataset1, its result is not that representive (not as well as the original or subdataset2) as this one contains only a quite small dataset. So we can still believe that they are not so extremely different.<br/>
<br/>
<br/>
<br/>